Galaxy Zoo: Morphological Classifications for 120,000 Galaxies in HST Legacy Imaging

### 2016-09-30 writeup/gz_hubble_data.tex

Classifications for GZH were weighted only if the classifier was logged
into the site under their username (this was encouraged, but not required for
participation). Classifications by participants who were not logged in were marked as
``Anonymous'' and receive the same $w=1$ weighting as the vast majority of
logged-in classifiers. \changes{Results from the Galaxy Zoo: CANDELS project (Simmons et~al., submitted)
show that the distributions of weights for anonymous and
logged-in users are similar, supporting the default technique of GZH.}
% We should also try to estimate how many anonymous classifiers there were, if possible. 
% Not clear that I can do that given the SQL data

@@

\changes{Figure~\ref{fig:sb_redshift} shows the distribution in $\mu$ and $z$ of
the artificially-redshifted \textsc{FERENGI} images compared to the genuine
\hst{} images. The full bivariate distributions differ due to a combination of
the detection limits of the \hst{} surveys, intrinsic rareness of bright $\mu$
galaxies in the SDSS volume, and evolution of the stellar populations. Since
the dependence of the debiasing correction is evaluated in separate bins, the
main concern is the existence of an overlap between the two sets of images;
this is the case for almost all of the \hst{} images, which have $\mu >
19$~mag~arcsec$^{-2}$.  The brightest end of the distribution, which has fewer
examples of comparable \textsc{FERENGI} images, correspond to galaxies for
which the debiased correction is expected to be minor.}

@@

\changes{\subsection{Effects of morphological debiasing}\label{ssec:zeta_results}}
The approach used in GZH for correcting the weighted classifications for
redshift bias rests on the assumption that the \emph{degree} of bias is a
function of the apparent size and brightness of the galaxy.
This is controlled by two types of parameters: \textbf{intrinsic} properties of
the galaxy itself, such as its physical diameter and luminosity, and
\textbf{extrinsic} properties, such as the distance (redshift) of the galaxy
and its relative orientation. The combination of all
such parameters forms a high-dimensional space, \changes{and there is no obvious technique
for separating these into individual effects}. Instead, the method used here employs
only two parameters as a simple model of the effect of redshift brightness on classification: the
surface brightness ($\mu$; intrinsic) and redshift ($z$;
extrinsic).

@@

For the ``lower-limit'' galaxies for which a single debiased \ffeatures{} value
cannot be confidently assigned, the \emph{range} of debiased values is
estimated and included as a data product. This uses the
\ferengi{} simulated data to analyse the range of intrinsic \ffeaturesrest{}
values for any given observed \ffeatures{} value, again as a function of
surface brightness and redshift. \changes{In each unclassifiable (shaded blue) $z$,$\mu$ bin, 
we compute the spread of
intrinsic values of \ffeaturesrest{} for five quantiles of observed \ffeatures{}
(corresponding to the ``clean'' thresholds used in prior GZ publications).
The range of intrinsic values for GZH is defined by the upper and lower 1~$\sigma$ limits,
enclosing the inner 68\% of the data. This range is represented by the orange bars in 
Figure}~\ref{fig:f_vs_f_zoom}. For any galaxy which
cannot be directly debiased by the $\zeta$ method, these ranges are used to
denote the upper and lower limits on the expected values \ffeaturesrest{} as a
function of the observed \ffeatures.

@@

\changes{\subsection{Correcting morphologies for classification bias}\label{ssec:zeta}}

For the ``correctable'' sample of galaxies, we observe a decline in the vote fraction
\ffeatures{} with increasing simulated redshift for each unique galaxy. 
We model this relationship \changes{for artificially-redshifted images} with a simple exponential
function which bounds $f_{\mu, z>0.3}$  between $f_{\mu, z=0.3}$ and 0:

\changes{
\begin{equation}
f_{\mu,z} = 1 - (1 - f_{\mu,z=0.3})e^{\frac{z-z_0}{\hat\zeta}}
\label{eqn:fzeta}
\end{equation}
}
%old equation
%We model the debiased vote fraction $f_{\mu,z}$ for an
%artificially-redshifted galaxy as:

@@

\noindent \changes{where $f_{\mu,z=0.3}$ is the vote fraction at the lowest redshift in
the artificially-redshifted sample ($z_0=0.3$). Here, $\zeta$ is a parameter that
controls the rate at which \ffeatures{} decreases with increasing redshift (and
that may depend on other galaxy properties). This function bounds the observed
vote fractions between $f_{\mu,z=0.3}$ and zero based on two assumptions: 1)
the vote fractions for featured galaxies decrease monotonically with increasing
redshift, therefore restricting vote fractions for a given galaxy to be less
than $f_{\mu,z=0.3}$ (which is almost always true of the data), and 2) the
vote fractions cannot be less than zero (which is always true).}

@@

\begin{figure}
\center
\includegraphics[width=0.5\textwidth]{figures/zeta_mu.pdf}
\caption{All fits for the \ferengi{} galaxies of the vote fraction dropoff
parameter $\zeta$ for \ffeatures{} as a function of surface brightness. This
includes only the simulated galaxies with a bounded range on the dropoff
\changes{($-10<\zeta<10$)} and sufficient points to fit each function (\changes{28~original
galaxies, each with varying images artificially redshifted in one to eight bins over a range from} $0.3\lesssim z_\mathrm{sim}\lesssim1.0$). 
}
\label{fig:zeta_mu}
\end{figure}

@@

\noindent where $\hat\zeta$ is the correction factor applied to each galaxy as
a function of surface brightness. The best-fit parameters to the linear fit
from least-squares optimization are $\zeta_0=0.50$, $\zeta_1=-0.03$. To make the
final debiased correction \changes{for the genuine \hst{} data, we apply a 
correction similar to Equation~\ref{eqn:fzeta}:}

@@

\begin{figure}
\center
\includegraphics[width=0.5\textwidth]{figures/correctable_fraction.pdf}
\caption{\changes{Histogram showing the fraction of galaxies that have a finite correction
for the debiased vote fractions \ffeaturesdebiased{} as a function of \ffeatures{}
and redshift. The parameter space for corrections is limited to $0.3 \leq z \leq 1.0$
due to the sampling of the parent SDSS galaxies and detectability in the \ferengi{} images.}
}
\label{fig:correctable_fraction}
\end{figure}

\changes{Figure~\ref{fig:correctable_fraction} shows the
distribution of the fraction of galaxies that are ultimately correctable as a function
of \ffeatures{} and redshift. The distributions for individual samples (AEGIS, COSMOS,
GEMS, GOODS) are individually very similar. We emphasize that a correction is more likely to be derived
and applied for galaxies at higher redshift and with higher weighted values of \ffeatures.
This has important consequences for selection of physically meaningful samples from GZH
(see Section~\ref{sec:cookbook}), meaning that comparative studies of galaxies
should use a threshold on the likelihood of features that evolves as a function of redshift.}

@@

In the \ferengi{} data, these numbers severely limit the amount of information
that can be extracted for the higher-tier questions. The debiasing technique
used (Section~\ref{ssec:zeta}) requires that at least 10~classifiers answer each
question for a galaxy with $z_\mathrm{sim}=0.3$ \emph{and} the corresponding
image at higher redshift. This requirement is (by default) met by all galaxies
for the smooth/features question. However, this is often \emph{not} met for
questions beyond Task~01. On average, $60\%\pm24\%$ of the galaxies do not have
sufficient data to measure a correction, as compared to 2.0\% achieved for
Task~01 (Table~\ref{tbl:ferengi_corrections}). \changes{This leaves the average
surface brightness/redshift bin 
with insufficient data points to confidently measure the change in vote fraction ($\lesssim10$~galaxies per bin). 
For these reasons, debiased vote fractions are only provided in the GZH catalogue for Task~01
(smooth/features). We suggest that use of morphological data for higher order tasks should instead use
the weighted vote fractions (see Section~\ref{sec:cookbook}).}

@@

We repeated all analyses comparing morphologies between GZH and external
\changes{publications} on the entire GZH dataset, rather than the volume-limited sample
($z<1.0,m<22.5$) discussed above. If galaxies at all redshifts and at the
original magnitude limit of $m<23.5$ are used, the correlations between all of
the automatic morphological populations and GZH decrease significantly.  The
exceptions are the GEMS galaxies, which are the only sample that was visually
classified and had a shallower magnitude cut on their catalogue.  Based on
these comparisons, we suggest that use of morphological catalogues with these
\hst{} images be limited to galaxies with $m_\mathrm{I|i|z}<22.5$; fainter
targets have been demonstrated to be less reliable, regardless of the method of
morphological classification being used.

@@

The data in the GZH catalogues have been extensively reduced and tested. A
dominant effect is the known bias against identifying disky and asymmetric
sub-structures at either low resolution or surface brightness. This can be the
result either of genuinely small or dim galaxies, or a perceived effect from
observing galaxies at further distances (higher redshift). To calibrate this
\emph{without} potentially overcorrecting for the genuine morphological
evolution of galaxies over cosmic time, the GZH project uses SDSS images of
low-redshift galaxies, processes them to appear as if they were at higher
redshift, and classifies them through the GZH interface in an identical
fashion. The resulting change in \ffeatures{} as a function of $z$ and $\mu$ is
applied as a multiplicative correction to the top-level vote fractions for
$\sim50\%$ of the GZH galaxies. \changes{However, any population studies using
GZH morphological data must use a combination of debiased vote fractions
and upper limits from the ``uncorrectable'' galaxies, due to the evolving nature
of the thresholds as a function of redshift.}

@@

Of the 11,157 galaxies in the GOODS-N and GOODS-S full depth sample, 4,460
are also in the shallow-depth sample. Figure~\ref{fig:shallow_vs_full} shows a
strong correlation between \ffeatures{} for both sets of images. \changes{The mean
change in \ffeatures{} from the shallow to full depth images
$f_\mathrm{features,full} - f_{features,shallow} \equiv \Delta f = -0.01$, with
a standard deviation of $\sigma = 0.18$.} While there is some variance in
$\Delta f$ in the whole sample, the change is usually small and not often
significant enough to change a morphological classification. Defining a clean
sample of disk galaxies as those with \fbest$>0.8$, elliptical galaxies as
those with $f_\mathrm{smooth,best}<0.2$, and intermediate as those in between,
75\% of the sample would not change morphology. Of the remaining 25\% that
would change morphology, only \changes{0.2\% (9~galaxies)}
drastically change morphology either from smooth to featured or vice versa, while the
rest transition to or from the ``intermediate'' morphology. Details can
be seen in Table~\ref{tbl:shallow_to_full_stats} and examples of images
representing the 9 possible changes (or lack of) in morphology are shown in
Figures~\ref{fig:shallow_smooth},\ref{fig:shallow_intermediate}, and
\ref{fig:shallow_featured}.

>>>>

Classifications for GZH were weighted only if the classifier was logged
into the site under their username (this was encouraged, but not required for
participation). Classifications by participants who were not logged in were marked as
``Anonymous'' and receive the same $w=1$ weighting as the vast majority of
logged-in classifiers. Results from the Galaxy Zoo: CANDELS project (Simmons et~al., submitted)
show that the distributions of weights for anonymous and
logged-in users are similar, supporting the default technique of GZH.
% We should also try to estimate how many anonymous classifiers there were, if possible. 
% Not clear that I can do that given the SQL data

@@

Figure~\ref{fig:sb_redshift} shows the distribution in $\mu$ and $z$ of
the artificially-redshifted \textsc{FERENGI} images compared to the genuine
\hst{} images. The full bivariate distributions differ due to a combination of
the detection limits of the \hst{} surveys, intrinsic rareness of bright $\mu$
galaxies in the SDSS volume, and evolution of the stellar populations. Since
the dependence of the debiasing correction is evaluated in separate bins, the
main concern is the existence of an overlap between the two sets of images;
this is the case for almost all of the \hst{} images, which have $\mu >
19$~mag~arcsec$^{-2}$.  The brightest end of the distribution, which has fewer
examples of comparable \textsc{FERENGI} images, correspond to galaxies for
which the debiased correction is expected to be minor.

@@

\subsection{Effects of morphological debiasing}\label{ssec:zeta_results}
The approach used in GZH for correcting the weighted classifications for
redshift bias rests on the assumption that the \emph{degree} of bias is a
function of the apparent size and brightness of the galaxy.
This is controlled by two types of parameters: \textbf{intrinsic} properties of
the galaxy itself, such as its physical diameter and luminosity, and
\textbf{extrinsic} properties, such as the distance (redshift) of the galaxy
and its relative orientation. The combination of all
such parameters forms a high-dimensional space, and there is no obvious technique
for separating these into individual effects. Instead, the method used here employs
only two parameters as a simple model of the effect of redshift brightness on classification: the
surface brightness ($\mu$; intrinsic) and redshift ($z$;
extrinsic).

@@

For the ``lower-limit'' galaxies for which a single debiased \ffeatures{} value
cannot be confidently assigned, the \emph{range} of debiased values is
estimated and included as a data product. This uses the
\ferengi{} simulated data to analyse the range of intrinsic \ffeaturesrest{}
values for any given observed \ffeatures{} value, again as a function of
surface brightness and redshift. In each unclassifiable (shaded blue) $z$,$\mu$ bin, 
we compute the spread of
intrinsic values of \ffeaturesrest{} for five quantiles of observed \ffeatures{}
(corresponding to the ``clean'' thresholds used in prior GZ publications).
The range of intrinsic values for GZH is defined by the upper and lower 1~$\sigma$ limits,
enclosing the inner 68\% of the data. This range is represented by the orange bars in 
Figure~\ref{fig:f_vs_f_zoom}. For any galaxy which
cannot be directly debiased by the $\zeta$ method, these ranges are used to
denote the upper and lower limits on the expected values \ffeaturesrest{} as a
function of the observed \ffeatures.

@@

\subsection{Correcting morphologies for classification bias}\label{ssec:zeta}

For the ``correctable'' sample of galaxies, we observe a decline in the vote fraction
\ffeatures{} with increasing simulated redshift for each unique galaxy. 
We model this relationship for artificially-redshifted images with a simple exponential
function which bounds $f_{\mu, z>0.3}$  between $f_{\mu, z=0.3}$ and 0:

\begin{equation}
f_{\mu,z} = 1 - (1 - f_{\mu,z=0.3})e^{\frac{z-z_0}{\hat\zeta}}
\label{eqn:fzeta}
\end{equation}
%old equation
%We model the debiased vote fraction $f_{\mu,z}$ for an
%artificially-redshifted galaxy as:

@@

\noindent where $f_{\mu,z=0.3}$ is the vote fraction at the lowest redshift in
the artificially-redshifted sample ($z_0=0.3$). Here, $\zeta$ is a parameter that
controls the rate at which \ffeatures{} decreases with increasing redshift (and
that may depend on other galaxy properties). This function bounds the observed
vote fractions between $f_{\mu,z=0.3}$ and zero based on two assumptions: 1)
the vote fractions for featured galaxies decrease monotonically with increasing
redshift, therefore restricting vote fractions for a given galaxy to be less
than $f_{\mu,z=0.3}$ (which is almost always true of the data), and 2) the
vote fractions cannot be less than zero (which is always true).

@@

\begin{figure}
\center
\includegraphics[width=0.5\textwidth]{figures/zeta_mu.pdf}
\caption{All fits for the \ferengi{} galaxies of the vote fraction dropoff
parameter $\zeta$ for \ffeatures{} as a function of surface brightness. This
includes only the simulated galaxies with a bounded range on the dropoff
($-10<\zeta<10$) and sufficient points to fit each function (28~original
galaxies, each with varying images artificially redshifted in one to eight bins over a range from $0.3\lesssim z_\mathrm{sim}\lesssim1.0$). 
}
\label{fig:zeta_mu}
\end{figure}

@@

\noindent where $\hat\zeta$ is the correction factor applied to each galaxy as
a function of surface brightness. The best-fit parameters to the linear fit
from least-squares optimization are $\zeta_0=0.50$, $\zeta_1=-0.03$. To make the
final debiased correction for the genuine \hst{} data, we apply a 
correction similar to Equation~\ref{eqn:fzeta}:

@@

\begin{figure}
\center
\includegraphics[width=0.5\textwidth]{figures/correctable_fraction.pdf}
\caption{Histogram showing the fraction of galaxies that have a finite correction
for the debiased vote fractions \ffeaturesdebiased{} as a function of \ffeatures{}
and redshift. The parameter space for corrections is limited to $0.3 \leq z \leq 1.0$
due to the sampling of the parent SDSS galaxies and detectability in the \ferengi{} images.
}
\label{fig:correctable_fraction}
\end{figure}

Figure~\ref{fig:correctable_fraction} shows the
distribution of the fraction of galaxies that are ultimately correctable as a function
of \ffeatures{} and redshift. The distributions for individual samples (AEGIS, COSMOS,
GEMS, GOODS) are individually very similar. We emphasize that a correction is more likely to be derived
and applied for galaxies at higher redshift and with higher weighted values of \ffeatures.
This has important consequences for selection of physically meaningful samples from GZH
(see Section~\ref{sec:cookbook}), meaning that comparative studies of galaxies
should use a threshold on the likelihood of features that evolves as a function of redshift.

@@

In the \ferengi{} data, these numbers severely limit the amount of information
that can be extracted for the higher-tier questions. The debiasing technique
used (Section~\ref{ssec:zeta}) requires that at least 10~classifiers answer each
question for a galaxy with $z_\mathrm{sim}=0.3$ \emph{and} the corresponding
image at higher redshift. This requirement is (by default) met by all galaxies
for the smooth/features question. However, this is often \emph{not} met for
questions beyond Task~01. On average, $60\%\pm24\%$ of the galaxies do not have
sufficient data to measure a correction, as compared to 2.0\% achieved for
Task~01 (Table~\ref{tbl:ferengi_corrections}). This leaves the average
surface brightness/redshift bin 
with insufficient data points to confidently measure the change in vote fraction ($\lesssim10$~galaxies per bin). 
For these reasons, debiased vote fractions are only provided in the GZH catalogue for Task~01
(smooth/features). We suggest that use of morphological data for higher order tasks should instead use
the weighted vote fractions (see Section~\ref{sec:cookbook}).

@@

We repeated all analyses comparing morphologies between GZH and external
publications on the entire GZH dataset, rather than the volume-limited sample
($z<1.0,m<22.5$) discussed above. If galaxies at all redshifts and at the
original magnitude limit of $m<23.5$ are used, the correlations between all of
the automatic morphological populations and GZH decrease significantly.  The
exceptions are the GEMS galaxies, which are the only sample that was visually
classified and had a shallower magnitude cut on their catalogue.  Based on
these comparisons, we suggest that use of morphological catalogues with these
\hst{} images be limited to galaxies with $m_\mathrm{I|i|z}<22.5$; fainter
targets have been demonstrated to be less reliable, regardless of the method of
morphological classification being used.

@@

The data in the GZH catalogues have been extensively reduced and tested. A
dominant effect is the known bias against identifying disky and asymmetric
sub-structures at either low resolution or surface brightness. This can be the
result either of genuinely small or dim galaxies, or a perceived effect from
observing galaxies at further distances (higher redshift). To calibrate this
\emph{without} potentially overcorrecting for the genuine morphological
evolution of galaxies over cosmic time, the GZH project uses SDSS images of
low-redshift galaxies, processes them to appear as if they were at higher
redshift, and classifies them through the GZH interface in an identical
fashion. The resulting change in \ffeatures{} as a function of $z$ and $\mu$ is
applied as a multiplicative correction to the top-level vote fractions for
$\sim50\%$ of the GZH galaxies. However, any population studies using
GZH morphological data must use a combination of debiased vote fractions
and upper limits from the ``uncorrectable'' galaxies, due to the evolving nature
of the thresholds as a function of redshift.

@@

Of the 11,157 galaxies in the GOODS-N and GOODS-S full depth sample, 4,460
are also in the shallow-depth sample. Figure~\ref{fig:shallow_vs_full} shows a
strong correlation between \ffeatures{} for both sets of images. The mean
change in \ffeatures{} from the shallow to full depth images
$f_\mathrm{features,full} - f_{features,shallow} \equiv \Delta f = -0.01$, with
a standard deviation of $\sigma = 0.18$. While there is some variance in
$\Delta f$ in the whole sample, the change is usually small and not often
significant enough to change a morphological classification. Defining a clean
sample of disk galaxies as those with \fbest$>0.8$, elliptical galaxies as
those with $f_\mathrm{smooth,best}<0.2$, and intermediate as those in between,
75\% of the sample would not change morphology. Of the remaining 25\% that
would change morphology, only 0.2\% (9~galaxies)
drastically change morphology either from smooth to featured or vice versa, while the
rest transition to or from the ``intermediate'' morphology. Details can
be seen in Table~\ref{tbl:shallow_to_full_stats} and examples of images
representing the 9 possible changes (or lack of) in morphology are shown in
Figures~\ref{fig:shallow_smooth},\ref{fig:shallow_intermediate}, and
\ref{fig:shallow_featured}.
